### 2017-11-04 pkg/model4you/inst/JORS/man_model4you.Rnw
working on comments from Torsten
git-svn-id: svn://svn.r-forge.r-project.org/svnroot/partykit@1675 255f4249-2263-4417-b1ea-8498779e27a7

Former-commit-id: 273888732638b6079c9c85f2c71efc5062aef546

@@

<<setup, echo=FALSE>>=
library("knitr")
knitr::opts_chunk$set(cache = TRUE, message = FALSE, fig.align = "center")

knit_hooks$set(inline = function(x) { 
  if (is.numeric(x)) return(knitr:::format_sci(x, 'latex')) 
  highr:::hi_latex(x) 
}) 
@

@@

To investigate whether the exam was fair, we asses whether the two groups
differ in terms of the percentage of correctly answered exam questions. This
can be done using a simple linear model regressing the percentage points of
correct anwers on the exam group. 
<<math_bmod>>=
bmod_math <- lm(pcorrect ~ group, data = MathExam)
@

The model can be visualised by plotting the estimated densities
(see Figure~\ref{fig:math_bmod_vis} (a)):
<<eval=FALSE>>=  
lm_plot(bmod_math)
@

As the plot is a \emph{ggplot2} graphic it can be maniulated with the usual
\emph{ggplot2} layers, e.g.\ with
<<eval=FALSE>>=
lm_plot(bmod_math) + labs(y = "Estimated Density", x = "% Correct")
@
or equivalently with 
<<eval=FALSE>>=
myplot <- function(model, ...) {
  lm_plot(model, ...) + 
    labs(y = "Estimated Density", x = "% Correct")
}
myplot(bmod_math)
@
The result is shown in Figure~\ref{fig:math_bmod_vis} (b).
<<math_bmod_vis, echo=FALSE, fig.cap = "Base model for Mathematics Exam data.", fig.subcap=c("Figure made with \\texttt{lm\\_plot()}.", "Figure made with manipulated \\texttt{lm\\_plot()}."), out.width="0.45\\textwidth", fig.height=3, fig.width=5>>=
lm_plot(bmod_math)
myplot <- function(model, ...) {
  lm_plot(model, ...) + 
    labs(y = "Estimated Density", x = "% Correct")
}
myplot(bmod_math)
@

A tree based on this model can be computed and visualised in only two lines of
code:
<<math_tree, fig.width = 13, fig.cap = "Personalised model tree for math exam.">>=  
tr_math <- pmtree(bmod_math, control = ctree_control(maxdepth = 2))
plot(tr_math, terminal_panel = node_pmterminal(tr_math, plotfun = myplot))
@

@@

ggplot(dpdat_math, aes(x = tests, y = group2)) +
  geom_point(alpha = 0.2, size = 1) + 
  geom_smooth(fill = NA, method = "loess") +
  theme(legend.position = "none")

ggplot(dpdat_math, aes(x = tests, y = group2, color = tests)) +
  geom_quasirandom(alpha = 0.5, size = 1) +
  theme(legend.position = "none")
@

@@

ggplot(dpdat_math, aes(x = attempt, y = group2, color = attempt)) +
  geom_quasirandom(alpha = 0.5) +
  theme(legend.position = "none")

ggplot(dpdat_math, aes(x = gender, y = group2, color = gender)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  theme(legend.position = "none")
@

@@

\section*{Implementation and architecture}
The R package \emph{model4you} is focused on ease of use and interpretability.
Users can take a simple model that they know and understand as basis and simply
plug it into \Sexpr{'pmtree()'} or \Sexpr{'pmforest()'} depending on whether
they want subgroup wise or personalised models. The basis for these
functionalities is provided by the \emph{partykit} package which is a widely
used R package for trees and forests. The \emph{model4you} package provides
wrappers for the well implemented and tested functions
\Sexpr{'partykit::ctree()'} and \Sexpr{'partykit::cforest()'} and extends the
functionalities to allow for the computation of personalised models and to 
improve usability and interpretability.

@@

% \section*{References}
\bibliographystyle{plain}
\bibliography{ref_model4you}

>>>>

\usepackage{booktabs}

%% Nice inline code
\makeatletter
\newcommand*{\textalltt}{}
\DeclareRobustCommand*{\textalltt}{%
  \begingroup
    \let\do\@makeother
    \dospecials
    \catcode`\\=\z@
    \catcode`\{=\@ne
    \catcode`\}=\tw@
    \verbatim@font\@noligs
    \@vobeyspaces
    \frenchspacing
    \@textalltt
}
\newcommand*{\@textalltt}[1]{%
    #1%
  \endgroup
}
\makeatother

@@

<<setup, echo=FALSE, message=FALSE>>=
library("knitr")
knitr::opts_chunk$set(cache = TRUE, message = FALSE, fig.align = "center")

# knit_hooks$set(inline = function(x) {
#   if (is.numeric(x)) return(knitr:::format_sci(x, 'latex'))
#   highr:::hi_latex(x)
# })

library("dplyr")
library("stringr")

knit_hooks$set(inline = function(x) { 
  if (is.numeric(x)) return(knitr:::format_sci(x, 'latex')) 
  highr:::hi_latex(x) %>%  str_replace("hlopt", "textalltt")
  }) 
@

@@

To investigate whether the exam was fair, we asses whether the two groups
differ in terms of the percentage of correctly answered exam questions. This
can be done using a simple linear model regressing the percentage points of
correct anwers on the exam group. 
<<math_bmod>>=
bmod_math <- lm(pcorrect ~ group, data = MathExam)
@
The estimtates and confidence intervals of this model can be computed via
<<>>=
cbind(estimate = coef(bmod_math), confint(bmod_math))
@
% The estimtates and confidence intervals of this model are shown in 
% Table~\ref{tab:bmod}.
% <<math_bmod_coef, results="asis", echo=FALSE>>=
% library("Hmisc")
% cf <- round(cbind(estimate = coef(bmod_math), confint(bmod_math)), 2)
% colnams <- dimnames(cf)[[2]]
% colnams <- gsub("\\%", "\\\\%", colnams)
% colnams[1] <- "$\\theta$"
% 
% latex(cf,
%       booktabs = TRUE,
%       rowlabel = "", file = "", 
%       colheads = colnams,
%       label = "tab:bmod", caption = "Estimate and confidence intervals for the base model."
%       )
% @
%% <TH> warum brauchen wir die Abb 2x? Bitte nur einmal richtig ohne die
%% gg-Freuerei. /TH>
The model can be visualised by plotting the estimated densities
(see Figure~\ref{fig:math_bmod_vis}):
<<math_bmod_vis, fig.height=2.5, fig.width=5, out.width="0.7\\textwidth", fig.cap = "Density estimates of base model for the Mathematics Exam data.">>=  
lm_plot(bmod_math)
@
Both the estimates and confidence intervals and the density curves suggest
that there is almost no difference between the two groups. But does this
really hold for all types of students?
% As the plot is a \emph{ggplot2} graphic it can be maniulated with the usual
% \emph{ggplot2} layers, e.g.\ with
% <<eval=FALSE>>=
% lm_plot(bmod_math) + labs(y = "Estimated Density", x = "% Correct")
% @
% or equivalently with 
% <<eval=FALSE>>=
% myplot <- function(model, ...) {
%   lm_plot(model, ...) + 
%     labs(y = "Estimated Density", x = "% Correct")
% }
% myplot(bmod_math)
% @
% The result is shown in Figure~\ref{fig:math_bmod_vis} (b).
% <<math_bmod_vis, echo=FALSE, fig.cap = "Base model for Mathematics Exam data.", fig.subcap=c("Figure made with \\texttt{lm\\_plot()}.", "Figure made with manipulated \\texttt{lm\\_plot()}."), out.width="0.45\\textwidth", fig.height=3, fig.width=5>>=
% lm_plot(bmod_math)
% myplot <- function(model, ...) {
%   lm_plot(model, ...) + 
%     labs(y = "Estimated Density", x = "% Correct")
% }
% myplot(bmod_math)
% @

A tree based on this model can be computed and visualised in only two lines of
code:
<<math_tree, fig.width = 13, fig.cap = "Personalised model tree for the Mathematics Exam datam.">>=  
tr_math <- pmtree(bmod_math, control = ctree_control(maxdepth = 2))
plot(tr_math, terminal_panel = node_pmterminal(tr_math, plotfun = lm_plot))
@

@@

ggplot(dpdat_math, aes(x = tests, y = group2)) +
  geom_point(alpha = 0.2, size = 1) + 
  geom_smooth(fill = NA, method = "loess") +
  theme(legend.position = "none") +
  ylab("individual estmated group effect")

ggplot(dpdat_math, aes(x = tests, y = group2, color = tests)) +
  geom_quasirandom(alpha = 0.5, size = 1) +
  theme(legend.position = "none") +
  ylab("individual estmated group effect")
@

@@

ggplot(dpdat_math, aes(x = attempt, y = group2, color = attempt)) +
  geom_quasirandom(alpha = 0.5) +
  theme(legend.position = "none") +
  ylab("individual estmated group effect")

ggplot(dpdat_math, aes(x = gender, y = group2, color = gender)) +
  geom_boxplot() +
  geom_quasirandom(alpha = 0.5) +
  theme(legend.position = "none") +
  ylab("individual estmated group effect")
@

@@

\section*{Implementation and architecture}
The R package \emph{model4you} is focused on ease of use and interpretability.
Users can take a simple model that they know and understand as basis and simply
plug it into \Sexpr{'pmtree()'} or \Sexpr{'pmforest()'} depending on whether
they want subgroup wise or personalised models. The basis for these
functionalities is provided by the \emph{partykit} 
package which is a widely
used R package for trees and forests \citep{hothorn_partykit_2015, hothorn_partykit_2017}. 
The \emph{model4you} package provides
wrappers for the well implemented and tested functions
\Sexpr{'partykit::ctree()'} and \Sexpr{'partykit::cforest()'} and extends the
functionalities to allow for the computation of personalised models and to 
improve usability and interpretability.

@@

% \section*{References}
\bibliographystyle{plainnat}
\bibliography{ref_model4you}
